---
title: "RTT Predation Study - Fish Analysis"
subtitle: "Fish Abundance GLMMs, Temporal Trends, and Predation Correlations"
author: "Allyson DeMerlis"
date: "`r Sys.Date()`"
output: html_document
---

# Overview
This script covers:
1. Fish abundance GLMMs (parrotfish and butterflyfish across regions/strata/time)
2. Kruskal-Wallis + Dunn's post-hoc tests for mean counts
3. Figure S4: Fish abundance over time by region and strata
4. Figure S5: Site-level fish abundance vs predation correlations (corrected for pseudoreplication)
5. Fish-predation GLMMs (corrected random effects structure)

---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(here)
library(knitr)
library(tidyverse)
library(lubridate)
library(readxl)
library(plotrix)
library(cowplot)
library(ggpubr)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(car)
library(broom.mixed)
library(dunn.test)
library(ragg)

i_am("Analysis_AllysonDeMerlis.Rproj")
opts_knit$set(root.dir = here())

# ── PNAS theme ────────────────────────────────────────────────────────────────
theme_pnas <- function() {
  theme_classic() +
  theme(
    text         = element_text(size = 8),
    axis.title   = element_text(size = 8),
    axis.text    = element_text(size = 8),
    legend.text  = element_text(size = 8),
    legend.title = element_text(size = 8),
    plot.title   = element_text(size = 8, face = "bold"),
    strip.text   = element_text(size = 8),
    axis.line    = element_line(linewidth = 0.4),
    axis.ticks   = element_line(linewidth = 0.4)
  )
}
theme_set(theme_pnas())


# ── Color palettes ────────────────────────────────────────────────────────────
species_colors <- c(
  "M. cavernosa" = "#CC5500",
  "O. faveolata" = "#5A8754",
  "P. clivosa"   = "#E1AD01"
)

fish_colors <- c("Parrotfish" = "blue", "Butterflyfish" = "magenta")
```

---

# Load and Tidy Data

```{r load data, include=FALSE}
outplant_monitoring <- read_xlsx(
  "input_files/Supplementary Table 1 - Coral Outplant and Fish Data.xlsx",
  sheet = "tidyCoral RT_Outplant_Corrected"
)

fish_survey <- read_xlsx(
  "input_files/Supplementary Table 1 - Coral Outplant and Fish Data.xlsx",
  sheet = "tidy Fish RT_DETAIL"
)

# ── Apply consistent region names ─────────────────────────────────────────────
recode_regions <- function(df) {
  df %>%
    mutate(Region_name = case_when(
      Region == 1 ~ "Martin/Palm Beach",
      Region == 2 ~ "Broward-Miami",
      Region == 3 ~ "Biscayne",
      Region == 4 ~ "Upper Keys",
      Region == 5 ~ "Middle Keys",
      Region == 6 ~ "Lower Keys"
    )) %>%
    mutate(Region_name = fct_relevel(
      Region_name,
      "Martin/Palm Beach", "Broward-Miami", "Biscayne",
      "Upper Keys", "Middle Keys", "Lower Keys"
    ))
}

outplant_monitoring <- recode_regions(outplant_monitoring)
fish_survey         <- recode_regions(fish_survey)

outplant_monitoring %>% 
  mutate(Region = as.factor(Region), Site = as.factor(Site)) -> outplant_monitoring

# ── Site characteristics (depth, strata) ──────────────────────────────────────
outplant_monitoring %>%
  mutate(Region = as.factor(Region), Site = as.factor(Site)) %>%
  dplyr::select(Region, Site, DEPTH, STRATA) %>%
  distinct() -> site_characteristics

# ── Join strata/depth into fish survey ────────────────────────────────────────
fish_survey %>%
  mutate(Region = as.factor(Region), Site = as.factor(Site)) %>%
  left_join(site_characteristics, by = c("Region", "Site")) %>%
  mutate(unique_site_code = paste0(Region, "_", Site)) -> fish_survey

# ── Tidy outplant monitoring dates ────────────────────────────────────────────
outplant_fish_monitoring_dates <- fish_survey %>%
  dplyr::select(MI_CORRECTED, Region, Site, Date) %>%
  distinct()

outplant_monitoring_with_dates <- left_join(
  outplant_monitoring, outplant_fish_monitoring_dates,
  by = c("MI_CORRECTED", "Region", "Site")
)

middle_date_by_MI <- outplant_monitoring_with_dates %>%
  dplyr::select(MI_CORRECTED, Date) %>%
  distinct() %>%
  arrange(Date) %>%
  group_by(MI_CORRECTED) %>%
  slice(ceiling(n() / 2)) %>%
  drop_na() %>%
  rename(Middle_MI_date = Date)

outplant_monitoring_days <- left_join(
  outplant_monitoring_with_dates, middle_date_by_MI, by = "MI_CORRECTED"
) %>%
  mutate(
    Date_Monitored = coalesce(Date, Middle_MI_date),
    Exact_Days     = as.numeric(as.Date(Date_Monitored) - as.Date("2021-05-06")),
    Days_MI        = as.numeric(as.Date(Middle_MI_date) - as.Date("2021-05-06"))
  )

# ── Build outplant_monitoring_days_survival_glm (needed for S5 and fish-predation GLMMs) ──
outplant_monitoring_days %>%
  arrange(Region, Site, ColonyNumber, MI_CORRECTED) %>%
  group_by(Region, Site, ColonyNumber) %>%
  mutate(
    Experienced_Predation = ifelse(
      (FishBites == "yes" | PRED_1 > 0) & Snails == "no", "1", "0"
    ),
    Cumulative_Predation = ifelse(
      cumsum(ifelse(Experienced_Predation == "1", 1, 0)) > 0, "1", "0"
    ),
    Total_Predation_Events = cumsum(ifelse(Experienced_Predation == "1", 1, 0))
  ) %>%
  ungroup() %>%
  mutate(
    Species = case_when(
      Species == "MCAV" ~ "M. cavernosa",
      Species == "OFAV" ~ "O. faveolata",
      Species == "PCLI" ~ "P. clivosa"
    ),
    Survival = case_when(
      STATUS_2 == "alive"               ~ 1,
      STATUS_2 %in% c("dead", "dead_m") ~ 0
    )
  ) %>%
  mutate(
    Species               = as.factor(Species),
    MI_CORRECTED          = as.numeric(MI_CORRECTED),
    Region                = as.factor(Region),
    Site                  = as.factor(Site),
    SourceNursery         = as.factor(SourceNursery),
    STRATA                = as.factor(STRATA),
    ColonyNumber          = as.factor(ColonyNumber),
    Experienced_Predation = as.factor(Experienced_Predation),
    Cumulative_Predation  = as.factor(Cumulative_Predation)
  ) %>%
  filter(!(STATUS_2 == "dead" & (Snails == "yes" | FishBites == "yes" | PRED_1 > 0))) ->
  outplant_monitoring_days_survival_glm

# ── Join fish survey into coral monitoring data ───────────────────────────────
outplant_monitoring_days_survival_glm %>%
  right_join(
    fish_survey %>% dplyr::select(!SURVEY_MONTH),
    by = c("MI_CORRECTED", "Region", "Region_name", "Site", "Date", "DEPTH", "STRATA")
  ) -> outplant_monitoring_days_survival_glm_fish

# ── Fish survey long format for plots/tests ───────────────────────────────────
fish_survey %>%
  pivot_longer(Parrotfish:Butterflyfish, names_to = "Fish", values_to = "Count") %>%
  mutate(Fish = factor(Fish, levels = c("Parrotfish", "Butterflyfish"))) ->
  fish_survey_long

```

---

# Fish Abundance GLMMs

## Parrotfish

```{r parrotfish glmm}
# Best model: Region*STRATA + MI_CORRECTED (AIC 3146.6)
# random effect: unique_site_code to account for repeated surveys at same site
# zero-inflation by region: Martin/Palm Beach has structural zeros

glm_parrotfish <- glmmTMB(
  Parrotfish ~ Region * STRATA + MI_CORRECTED + (1|unique_site_code),
  family     = nbinom2,
  ziformula  = ~ Region,
  data       = fish_survey
)
summary(glm_parrotfish)

sim_res_parrotfish <- simulateResiduals(glm_parrotfish)
plot(sim_res_parrotfish)

glm_parrotfish_anova <- Anova(glm_parrotfish, type = "II")
print(glm_parrotfish_anova)

# Export for SI
as.data.frame(glm_parrotfish_anova) %>%
  rownames_to_column("Term") %>%
  write_csv("SI_tables/parrotfish_glmm_anova.csv")

broom.mixed::tidy(glm_parrotfish, effects = "fixed") %>%
  write_csv("SI_tables/parrotfish_glmm_fixed_effects.csv")

# Regional pairwise comparisons
emm_parrotfish_region <- emmeans(glm_parrotfish, ~ Region, type = "response")
pairs_parrotfish_region <- as.data.frame(pairs(emm_parrotfish_region))
print(pairs_parrotfish_region)
write_csv(pairs_parrotfish_region, "SI_tables/parrotfish_emmeans_region.csv")

# Region x Strata pairwise
emm_parrotfish_strata <- emmeans(glm_parrotfish, ~ Region * STRATA, type = "response")
pairs_parrotfish_strata <- as.data.frame(pairs(emm_parrotfish_strata))
write_csv(pairs_parrotfish_strata, "SI_tables/parrotfish_emmeans_region_strata.csv")
```

## Butterflyfish

```{r butterflyfish glmm}
# Best model: Region*STRATA + MI_CORRECTED (AIC 1524.5)

glm_butterflyfish <- glmmTMB(
  Butterflyfish ~ Region * STRATA + MI_CORRECTED + (1|unique_site_code),
  family    = nbinom2,
  ziformula = ~ Region,
  data      = fish_survey
)
summary(glm_butterflyfish)

sim_res_butterflyfish <- simulateResiduals(glm_butterflyfish)
plot(sim_res_butterflyfish)

glm_butterflyfish_anova <- Anova(glm_butterflyfish, type = "II")
print(glm_butterflyfish_anova)

as.data.frame(glm_butterflyfish_anova) %>%
  rownames_to_column("Term") %>%
  write_csv("SI_tables/butterflyfish_glmm_anova.csv")

broom.mixed::tidy(glm_butterflyfish, effects = "fixed") %>%
  write_csv("SI_tables/butterflyfish_glmm_fixed_effects.csv")

emm_butterflyfish_region <- emmeans(glm_butterflyfish, ~ Region, type = "response")
pairs_butterflyfish_region <- as.data.frame(pairs(emm_butterflyfish_region))
print(pairs_butterflyfish_region)
write_csv(pairs_butterflyfish_region, "SI_tables/butterflyfish_emmeans_region.csv")

emm_butterflyfish_strata <- emmeans(glm_butterflyfish, ~ Region * STRATA, type = "response")
pairs_butterflyfish_strata <- as.data.frame(pairs(emm_butterflyfish_strata))
write_csv(pairs_butterflyfish_strata, "SI_tables/butterflyfish_emmeans_region_strata.csv")
```

---

# Kruskal-Wallis + Dunn's Post-Hoc Tests

```{r kruskal wallis regional differences}
# Parrotfish: regional differences in mean counts across all time points
kruskal.test(Count ~ Region, data = filter(fish_survey_long, Fish == "Parrotfish"))

dt_parrotfish <- dunn.test(
  fish_survey_long$Count[fish_survey_long$Fish == "Parrotfish"],
  fish_survey_long$Region[fish_survey_long$Fish == "Parrotfish"],
  method = "holm", kw = TRUE, altp = TRUE
)

results_dt_parrotfish <- data.frame(
  comparison = dt_parrotfish$comparisons,
  Z          = round(dt_parrotfish$Z, 3),
  p_unadj    = signif(dt_parrotfish$altP, 3),
  p_adj      = signif(dt_parrotfish$altP.adjusted, 3)
)
print(results_dt_parrotfish)
write_csv(results_dt_parrotfish, "SI_tables/parrotfish_dunn_region.csv")

# Butterflyfish
kruskal.test(Count ~ Region, data = filter(fish_survey_long, Fish == "Butterflyfish"))

dt_butterflyfish <- dunn.test(
  fish_survey_long$Count[fish_survey_long$Fish == "Butterflyfish"],
  fish_survey_long$Region[fish_survey_long$Fish == "Butterflyfish"],
  method = "holm", kw = TRUE, altp = TRUE
)

results_dt_butterflyfish <- data.frame(
  comparison = dt_butterflyfish$comparisons,
  Z          = round(dt_butterflyfish$Z, 3),
  p_unadj    = signif(dt_butterflyfish$altP, 3),
  p_adj      = signif(dt_butterflyfish$altP.adjusted, 3)
)
print(results_dt_butterflyfish)
write_csv(results_dt_butterflyfish, "SI_tables/butterflyfish_dunn_region.csv")

# Region x Strata pairwise (Kruskal-Wallis)
fish_survey_long %>%
  mutate(group = paste0(Region, "-", STRATA)) -> fish_counts_long_strata

kruskal.test(Count ~ group, data = filter(fish_counts_long_strata, Fish == "Parrotfish"))
kruskal.test(Count ~ group, data = filter(fish_counts_long_strata, Fish == "Butterflyfish"))
```

---

# Figure S3: Fish Abundance Over Time by Region and Strata

```{r fig S3}
# Summary: mean ± SE per region x strata x date
fish_survey_summary <- fish_survey %>%
  pivot_longer(Parrotfish:Butterflyfish, names_to = "Fish", values_to = "Count") %>%
  mutate(Fish = factor(Fish, levels = c("Parrotfish", "Butterflyfish"))) %>%
  group_by(Region_name, STRATA, Date, Fish) %>%
  summarise(
    mean_count = mean(Count, na.rm = TRUE),
    se_count   = std.error(Count),
    .groups    = "drop"
  )

p1_S3<-fish_survey %>% 
  pivot_longer(Parrotfish:Butterflyfish, names_to = "Fish", values_to = "Count") %>% 
  mutate(SURVEY_MONTH = as.Date(SURVEY_MONTH)) %>% 
  group_by(Region_name, Fish) %>% 
  mutate(mean_count = mean(Count), se_count = std.error(Count)) %>% 
  distinct() %>% 
  ggplot(., aes(x=Region_name, y = mean_count, color = Fish)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_count-se_count, ymax = mean_count+se_count), width = 0.1) +
  scale_color_manual(values = c("magenta", "blue")) +
  labs(x="Region", y = "Fish Abundance")  +
  theme_pnas() +
  theme(legend.position = "none")

p2_S3 <- fish_survey %>% 
  pivot_longer(Parrotfish:Butterflyfish, names_to = "Fish", values_to = "Count") %>% 
  group_by(Region, SURVEY_MONTH, Region_name, Fish, STRATA) %>% 
  mutate(mean_count = mean(Count), se_count = std.error(Count)) %>% 
  distinct() %>%
  ggplot(., aes(x=SURVEY_MONTH, y = mean_count, color = Fish)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_count-se_count, ymax = mean_count+se_count), width = 0.2) +
  facet_wrap(~Region_name*STRATA, ncol = 2) +
  scale_color_manual(values = c("magenta", "blue")) +
  theme_classic() +
    scale_x_date(
    date_breaks = "4 month",
    date_labels = "%b %Y",
  ) +
  labs(x="Date", y = "Fish Abundance")  +
  theme_pnas() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

plot_grid(p1_S3, p2_S3, ncol = 1, rel_heights = c(0.5, 2), labels = c("A", "B")) -> p_S3

ggsave("plots/Manuscript Figures/FigS3_fish_abundance_over_time.pdf",
       plot   = p_S3,
       width  = 16.5, height = 22, units = "cm",
       dpi    = 600)
```

---

# Figure S4: Fish Abundance vs Predation — Site-Level

```{r fig S4 site level aggregation}
# Aggregate to site level — n=24 independent observations
# Fish means across full study
fish_site_means <- outplant_monitoring_days_survival_glm_fish %>%
  filter(!(MI_CORRECTED == 2 & Region == 3)) %>%
  group_by(Region_name, Site) %>%
  summarise(
    mean_parrotfish    = mean(Parrotfish,    na.rm = TRUE),
    mean_butterflyfish = mean(Butterflyfish, na.rm = TRUE),
    .groups = "drop"
  )

# Cumulative predation prevalence at final time point, averaged across species
cumul_prev_final <- outplant_monitoring_days_survival_glm %>%
  filter(MI_CORRECTED == 25) %>%
  group_by(Region_name, Site) %>%
  summarise(
    cumul_pred_prevalence = mean(Cumulative_Predation == "1", na.rm = TRUE),
    .groups = "drop"
  )

# Join
site_level <- left_join(fish_site_means, cumul_prev_final,
                         by = c("Region_name", "Site"))

# Spearman correlations — cumulative prevalence
cor_parrot_cumul <- cor.test(site_level$mean_parrotfish,
                              site_level$cumul_pred_prevalence,
                              method = "spearman")
cor_butter_cumul <- cor.test(site_level$mean_butterflyfish,
                              site_level$cumul_pred_prevalence,
                              method = "spearman")

cat("Parrotfish vs cumul prevalence: rho =", round(cor_parrot_cumul$estimate, 3),
    "  p =", round(cor_parrot_cumul$p.value, 3), "\n")
cat("Butterflyfish vs cumul prevalence: rho =", round(cor_butter_cumul$estimate, 3),
    "  p =", round(cor_butter_cumul$p.value, 3), "\n")

# Updated panels

pA <- ggplot(site_level, aes(x = mean_parrotfish, y = cumul_pred_prevalence)) +
  geom_point(aes(color = Region_name), size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "grey40",
              linewidth = 0.5, linetype = "dashed") +
  annotate("text",
           x = Inf, y = -Inf,
           hjust = 1.1, vjust = -0.5,
           label = paste0("Spearman R = ", round(cor_parrot_cumul$estimate, 2),
                          ", p = ", round(cor_parrot_cumul$p.value, 3)),
           size = 2.5, fontface = "italic") +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(x = "Mean parrotfish abundance",
       y = "Cumulative predation\nprevalence",
       color = NULL) +
  theme_pnas() +
  theme(legend.position = "none")

pB <- ggplot(site_level, aes(x = mean_butterflyfish, y = cumul_pred_prevalence)) +
  geom_point(aes(color = Region_name), size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "grey40",
              linewidth = 0.5, linetype = "dashed") +
   annotate("text",
           x = Inf, y = -Inf,
           hjust = 1.1, vjust = -0.5,
           label = paste0("Spearman R = ", round(cor_butter_cumul$estimate, 2),
                          ", p = ", round(cor_butter_cumul$p.value, 3)),
           size = 2.5, fontface = "italic") +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(x = "Mean butterflyfish abundance",
       y = "Cumulative predation\nprevalence",
       color = NULL) +
  theme_pnas()

legend_S4 <- get_legend(pB + theme(legend.position = "right"))

p_S4 <- plot_grid(
  pA,
  pB + theme(legend.position = "none"),
  legend_S4,
  ncol       = 3,
  labels     = c("A", "B", ""),
  rel_widths = c(1, 1, 0.5)
)

ggsave("plots/Manuscript Figures/FigS4_fish_abundance_vs_predation_sitelevel.pdf",
       plot   = p_S4,
       width  = 17.8, height = 8, units = "cm",
       dpi    = 600)
```


